---
title: "covid 19"
author: "Dr K"
date: "5/2/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```

```{r eval=FALSE}
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

```{r eval=FALSE}
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))
```



### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$ There is no correlation between sex and dying.

$H_A$ There is a correlation between sex and dying.

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

By looking at the ggplot bar chart that we created, we can observe that there are more males dead that females. The reason that this might have happened is because there were more females than males being observed. But even thought there were more female observed, when we look at the count we can see that the total amount of dead were mainly in the male side. about 65% of the male died and 38% of female died. 

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
Looking a the second table we can see that there is 1.5% of female end up dead and 3% of the males end up dead. This showsus that the percentage of dead in men is double times than female. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

From the Fisher squared test we see that men were 2.3 times more likely to die than females. The chances of getting this results if there realy was  no diference based on the p-value of 0.0009, this is a very small chance.

## Conclusions

In conclusion at the begging of this observation between females and males dying of covid-19 we can see in the ggplot that there were more woman being observed than male, however the bar chart also showed us that more males were dying than females. As we run our numeric result we can see that 1.5% percent of females died and 3% of male died. the death rate of males was twice higher  than females. The fisher square test showed us the p- value of 0.0009. being it such a small number there was no chance of this happening. 

